# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
```{r}
#data <- read.csv(unzip("activity.zip"))
data <- read.csv("activity.csv")
```

Remove the NA values
```{r}
data.noNA <- data[!is.na(data$steps),]
# remove unused factors
data.noNA$date <- factor(data.noNA$date)
```


## What is mean total number of steps taken per day?
Sum the steps for each day, then calculate the mean and median.
```{r}
total.steps.by.day <- tapply(data.noNA$steps, data.noNA$date, sum)
hist(total.steps.by.day, main="Histogram of Total Number of Steps per Day", xlab="")
mean(total.steps.by.day[!is.na(total.steps.by.day)])
```


## What is the average daily activity pattern?
Take the average number of steps by the interval across days.
```{r}
steps.by.daily.interval <- aggregate(steps ~ interval, data.noNA, FUN="mean")
with(steps.by.daily.interval, plot(interval, steps, type="l", 
                                   main="Average Daily Activity Pattern", 
                                   xlab="Interval (5 minute increment)",
                                   ylab="Average Number of Steps"))
```
Interval with the maximum number of steps on average across the day.
```{r}
steps.by.daily.interval[steps.by.daily.interval$steps == max(steps.by.daily.interval$steps),1]
```

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
